Fixing Vulnerabilities Hinders Maintainability

### 2019-02-03 paper/ICPC19/icpc19.tex
figura 5 fixed + example corrected

BCH compiles the ISO/IEC 25010 into a set of 10 guidelines that are the
base of BCH's evaluation. Two examples are the
\emph{Write Simple Units of Code} based on the McCabe Complexity~\cite{1702388}
and \emph{Keep your Codebase Small} based on the idea that smaller codebases are
easier to maintain and less prone to defects~\cite{Visser:2016:OREILLY}. Compliance to the guidelines is derived from the SIG's
industry benchmark which, according to official sources, consists of over 8
billion lines of code in more than 180 different technologies. SIG analyzes
around 15 million lines of code, per week.

@@

As an example, consider the refactoring of the Datagram Transport Layer Security
(DTLS) protocol implementation in OpenSSL\footnote{OpenSSL is a tookit that
contains open-source implementations of the SSL and TLS cryptographic
protocols.} to address a vulnerability where an hacker could cause a Denial-of-Service
(DoS) attack by crafting DTLS handshake messages to trigger memory allocations
corresponding to large length values. This vulnerability is listed at the Common
Vulnerabilities and Exposures dictionary as CVE-2014-3506\footnote{CVE-2014-3506
details available at http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2014-3506
(Accessed on \today{})}. It is amongst the vulnerabilities studied in our
evaluation. The following snippet presents the changes performend on the
\emph{ssl/d1\_both.c} file\footnote{CVE-2014-3506 fix available  at
https://github.com/openssl/openssl/commit/\\1250f12613b61758675848f6600ebd914ccd7636
(Accessed on \today{})} by the OpenSLL developers to fix the vulnerability.

@@

In the example (Listing~\ref{lst:vuln}), all functions receive as input the
\textbf{handshake} represented by \texttt{s}. Every SSL/TLS connection begins
with a handshake which is reponsible by the negotiation between two parties. The
handshake determines what cryptographic algorithms will be used to encrypt
communications, verify the server and establish a secure connection before
beginning data transfer. In the \texttt{dtls1\_reassemble\_fragment} function
the value of \texttt{msg\_hdr-$>$frag\_off+frag\_len} is checked against the
maximum handshake message size, but \texttt{msg\_len} is allocated to the
fragment buffer in line $37$. Thus, if the fragment was within the allowed size,
the pending handshake message could use up to $18$MB ($16$MB+$2$MB).
According to the message commit, $10$ handshake messages are allowed. Hence,
the attacker could potentially use $180$MB per DTLS connection.

The code changes performed to fix the vulnerability were the following:

~\ref{lst:func1} \texttt{dtls1\_max\_handshake\_message\_len} function was inserted to limit the number of bytes allowed in a DTLS handshake message,
\texttt{s}. Checks if the maximum certificate list size
(\texttt{s-$>$max\_cert\_list}) requires more length than the maximum length
(\texttt{max\_len}). If that condition holds, the function returns the size of
the certificate list, otherwise it returns the maximum length. This function replaces lines $18$-$21$.

~\ref{lst:func5} and~\ref{lst:func4} add the check to ensure that the message
length is never higher than the maxiumum allowed by the DTSL protocol in
different functions.

Although limiting the size of a handshake message seems an elementary problem to
repair, the example shows that there was a significant amount of changes
performed in the codebase which produced a negative impact in the project
maintainability. Even though the developer tries to simplify the codebase with
the creation of a smaller unit \texttt{dtls1\_max\_handshake\_message\_len} (\ref{lst:func1}), it
also disrupts $2$ of the guidelines proposed by the Software Improvement Group
(SIG) for building maintainable software~\cite{Visser:2016:OREILLY}. It adds a
new branch point to the \texttt{dtls1\_process\_out\_of\_seq\_message} unit
(\ref{lst:func5}) increasing the cyclomatic complexity and breaking the
\emph{Write Simple Units of Code} guideline. Moreover, it increases the codebase
size in $4$ lines of code ($13$ lines added, $9$ lines deleted) which contributes
to a potential violation of the
\emph{Keep Your Codebase Small} guideline. Hence, we have observed that the
impact of this refactoring on the OpenSSL maintainability was negative.

In this paper, our concern is to study whether, while improving software
security, developers are also negatively impacting the maintainability
of their software application. To answer the following two
research questions, we use a dataset of security refactorings~\cite{Reis:2017:IJSSE} to measure the
impact of the changes in the maintainability of open-source software.

@@

	\item \textbf{Broken Authentication \& Session Management.} Uncorrectly
	implemented functionalities related to authentication and session management,
	allowing crackers to gain access to session tokens, passwords, keys and other
	sensitive data. \textit{Based on the OWASP Top 10 of 2013.}

	\item \textbf{Cross-Site Request Forgery.} Poor session tokens generation and
	management usually allow crackers to send forged HTTP requests including
	authentication information from the victim to vulnerable web application.
	\textit{Based on the OWASP Top 10 of 2013.}

	\item \textbf{Denial-of-Service} Security flaws that can allow a cracker to
	flood or crashing services. This attacks usually occur when the system
	receives much traffic causing it to slow down and eventually stop. For
	example, flaws allowing crackers to trigger memory allocations corresponding
	to large length values (Listing~\ref{lst:vuln}).

@@

	\item \textbf{Cross-Site Scripting.} Lack of proper validation or escaping
	allow crackers to submit untrusted data to web browsers through malicious
	scripts that can hijack the user sessions or redirect the user to malicious
	sites. \textit{Based on the OWASP Top 10 of 2017.}

	\item \textbf{Injection.} When developers do not keep untrusted data separate
	from commands and queries. If a cracker sends a string that exploits the
	syntax of the interpreter, then an injection attack is possible (e.g., SQL and
	LDAP injection). \textit{Based on the OWASP Top 10 of 2017.}

>>>>

BCH compiles the ISO/IEC 25010 into a set of 10 guidelines that are the
base of BCH's evaluation. Two examples are the
\emph{Write Simple Units of Code} based on the McCabe Complexity~\cite{1702388}
and \emph{Keep your Codebase Small} based on the idea that smaller codebases are
easier to maintain and less prone to defects~\cite{Visser:2016:OREILLY}. 
% Compliance to the guidelines is derived from the SIG's
% industry benchmark which, according to official sources, consists of over 8
% billion lines of code in more than 180 different technologies. SIG analyzes
% around 15 million lines of code, per week.

@@

As an example, consider the refactoring of the Datagram Transport Layer Security
(DTLS) protocol implementation in OpenSSL\footnote{OpenSSL is a tookit that
contains open-source implementations of the SSL and TLS cryptographic
protocols.} to address a vulnerability where an attacker could cause a Denial-of-Service
(DoS) attack by crafting DTLS handshake messages to trigger memory allocations
corresponding to large length values. This vulnerability is listed at the Common
Vulnerabilities and Exposures dictionary as CVE-2014-3506\footnote{CVE-2014-3506
details available at http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2014-3506
(Accessed on \today{})}. It is amongst the vulnerabilities studied in our
evaluation. The following snippet presents the changes performend on the
\emph{ssl/d1\_both.c} file\footnote{CVE-2014-3506 fix available  at
https://github.com/openssl/openssl/commit/\\1250f12613b61758675848f6600ebd914ccd7636
(Accessed on \today{})} by the OpenSLL developers to patch the vulnerability.

@@

Every SSL/TLS connection begins with a handshake which is responsible by the negotiation between two parties. 
DTLS provides a mechanism for fragmenting a handshake message over a number of packets, each of which can be 
transmitted separately. In the example (Listing~\ref{lst:vuln}), all functions receive as input the
SSL session, \texttt{s}. In the \texttt{dtls1\_reassemble\_fragment} function the size of the fragment \texttt{msg\_hdr-$>$frag\_off+frag\_len} is checked against the handshake message size \texttt{msg\_hdr->msg\_len} 
(line 13) and then against the maximum handshake message size allowed by DTLS1 for \texttt{s} (line 16-21). 
However, it is not verified if \texttt{msg\_hdr->msg\_len} is lower than the allowed by DTLS. Thus, in line 37,
when \texttt{msg\_len} is allocated to the fragment buffer, more memory than the allowed might be allocated. According to the message commit, $10$ handshake messages are allowed per DTLS connection. Hence, a remote attacker could leverage this issue to trigger
memory allocations corresponding to large length values and lead the system to memory exhaustion that could end
in a DoS attack. The code changes performed to fix the vulnerability were the following:

~\ref{lst:func1} \texttt{dtls1\_max\_handshake\_message\_len} function was inserted to return the number of 
maximum number of bytes allowed in a DTLS handshake message for \texttt{s}. The mininum is $16$KB, but may be greater 
if the maximum certificate list size requires it which is set by the user. More specifically, it checks if the 
maximum certificate list size
(\texttt{s-$>$max\_cert\_list}) requires more length than the maximum length
(\texttt{max\_len}). If that condition holds, the function returns the size of
the certificate list, otherwise it returns the maximum length. This function replaces lines $18$-$21$.

~\ref{lst:func5} and~\ref{lst:func4} add the check to different functions to ensure that the message 
and the fragment length are never higher than the maximum allowed by DTSL. However, \ref{lst:func5} is
the responsible for patching the vulnerability.

Although checking the maximum size of a handshake message seems an elementary problem to
repair, the example shows that there was a significant amount of changes
performed in the codebase which produced a negative impact in the project
maintainability. Even though the developer tries to simplify the codebase with
the creation of a smaller unit \texttt{dtls1\_max\_handshake\_message\_len} (\ref{lst:func1}), it
also disrupts $2$ of the guidelines proposed by the Software Improvement Group
(SIG) for building maintainable software~\cite{Visser:2016:OREILLY}. It adds a
new branch point to the \texttt{dtls1\_process\_out\_of\_seq\_message} unit
(\ref{lst:func5}) increasing the cyclomatic complexity and breaking the
\emph{Write Simple Units of Code} guideline. Moreover, it increases the codebase
size in $4$ lines of code ($13$ lines added, $9$ lines deleted) which contributes
to a potential violation of the
\emph{Keep Your Codebase Small} guideline. Hence, we have observed that the
impact of this refactoring on the OpenSSL maintainability was negative.

In this paper, our concern is to study whether, while improving software
security, developers are also negatively impacting the maintainability
of their software applications. To answer the following two
research questions, we use a dataset of security refactorings~\cite{Reis:2017:IJSSE} to measure the
impact of the changes in the maintainability of open-source software.

@@

	\item \textbf{Broken Authentication \& Session Management.} Uncorrectly
	implemented functionalities related to authentication and session management,
	allowing attacker to gain access to session tokens, passwords, keys and other
	sensitive data. \textit{Based on the OWASP Top 10 of 2013.}

	\item \textbf{Cross-Site Request Forgery.} Poor session tokens generation and
	management usually allow attackers to send forged HTTP requests including
	authentication information from the victim to vulnerable web application.
	\textit{Based on the OWASP Top 10 of 2013.}

	\item \textbf{Denial-of-Service} Security flaws that can allow a attacker to
	flood or crashing services. This attacks usually occur when the system
	receives much traffic causing it to slow down and eventually stop. For
	example, flaws allowing attackers to trigger memory allocations corresponding
	to large length values (Listing~\ref{lst:vuln}).

@@

	\item \textbf{Cross-Site Scripting.} Lack of proper validation or escaping
	allow attackers to submit untrusted data to web browsers through malicious
	scripts that can hijack the user sessions or redirect the user to malicious
	sites. \textit{Based on the OWASP Top 10 of 2017.}

	\item \textbf{Injection.} When developers do not keep untrusted data separate
	from commands and queries. If a attacker sends a string that exploits the
	syntax of the interpreter, then an injection attack is possible (e.g., SQL and
	LDAP injection). \textit{Based on the OWASP Top 10 of 2017.}
